---
title: "Virus diversity in insects based on virushost database"
output:
  html_document: 
    toc: yes
    number_sections: yes
  html_notebook: default
  pdf_document: default
---

```{r, echo=FALSE, message = FALSE}
library("data.table")
library("taxize")
library("myTAI")
library(plyr)
library(ggvenn)
```

# import virus-host db from https://www.genome.jp/virushostdb/

```{r}
virus_host=as.data.frame(fread("~/Downloads/virushostdb.tsv"))
dim(virus_host)
head(virus_host)

# subset to insects
virus_host_insects=subset(x = virus_host, 
                          subset = grepl(pattern = "Insecta", virus_host$`host lineage`))
dim(virus_host_insects)
head(virus_host_insects)
```

Remove redundancy if any
```{r}
virus_host_insects=virus_host_insects[-which(duplicated(virus_host_insects$`virus tax id`)),]
dim(virus_host_insects)
```

One line is one virus species

# obtain full taxo info for viruses (remote, may be interrupted)

```{r, echo=TRUE, message = FALSE}
full_taxo_viruses=taxize::classification(virus_host_insects$`virus tax id`,db="ncbi")
```

```{r}
 length(full_taxo_viruses)
 # rearrange ranks information to produce a nice table
 full_taxo_viruses2=lapply(full_taxo_viruses, FUN = function(x){
   taxid=last(x$id)
   t=data.frame(taxid, t(x[, c("name")]))
   colnames(t)=c("taxid", x$rank)
   return(t)
 })
 full_taxo_viruses_df=do.call(rbind.fill, full_taxo_viruses2)
head(full_taxo_viruses_df)
```

#merge virus taxo info with virus-host association
```{r}
 virus_host_insects2=merge(x=virus_host_insects, 
                           y=full_taxo_viruses_df, 
                           by.x="virus tax id", 
                           by.y="taxid", 
                           all.x=TRUE, 
                           all.y=FALSE)
 dim(virus_host_insects2)
 head(virus_host_insects2)

```

# import ictv taxo table

```{r}
ictv=as.data.frame(fread("~/Downloads/ICTV_Master_Species_List_2022_MSL38.v1.txt"))
dim(ictv)
```

```{r}
unique(ictv$`Genome Composition`)
```

Identify families for each genomic structure

```{r}
# identify families with dsRNA genomic structure
dsRNA_fams=unique(ictv$Family[grep("dsRNA", ictv$`Genome Composition`)])
# remove an empty
dsRNA_fams=dsRNA_fams[-which(dsRNA_fams=="")]
# identify families with dsRNA genomic structure
dsDNA_fams=unique(ictv$Family[grep("dsDNA", ictv$`Genome Composition`)])
# remove an empty
dsDNA_fams=dsDNA_fams[-which(dsDNA_fams=="")]
# identify families with ssRNA genomic structure
ssRNA_fams=unique(ictv$Family[grep("ssRNA", ictv$`Genome Composition`)])
# remove an empty
ssRNA_fams=ssRNA_fams[-which(ssRNA_fams=="")]
# identify families with ssDNA genomic structure
ssDNA_fams=unique(ictv$Family[grep("ssDNA", ictv$`Genome Composition`)])

```

Check that there is no overlap:
```{r}

library(ggvenn)

x = list(ssDNA= ssDNA_fams, ssRNA = ssRNA_fams,dsDNA= dsDNA_fams, dsRNA= dsRNA_fams)
ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF"),
  stroke_size = 0.5, set_name_size = 4
)

```

There is 1 !

```{r}
intersect(ssDNA_fams, dsDNA_fams)
```

Never mind, this is an archae virus family.

# Add genomic structure information into the table

```{r}
virus_host_insects2$genomic_struture=factor(x = NA,levels = c("dsRNA", "dsDNA", "ssRNA", "ssDNA"))
virus_host_insects2$genomic_struture[virus_host_insects2$family %in% dsRNA_fams]="dsRNA"
virus_host_insects2$genomic_struture[virus_host_insects2$family %in% dsDNA_fams]="dsDNA"
virus_host_insects2$genomic_struture[virus_host_insects2$family %in% ssRNA_fams]="ssRNA"
virus_host_insects2$genomic_struture[virus_host_insects2$family %in% ssDNA_fams]="ssDNA"

head(virus_host_insects2)

```
# Plots and tables

## subset to major orders
```{r}
# subset to Hymenoptera
virus_host_insects2_hymeno=virus_host_insects2[grep("Hymeno", virus_host_insects2$`host lineage`),]
# subset to Diptera
virus_host_insects2_diptera=virus_host_insects2[grep("Diptera", virus_host_insects2$`host lineage`),]
# subset to Coleoptera
virus_host_insects2_coleo=virus_host_insects2[grep("Coleoptera", virus_host_insects2$`host lineage`),]
# subset to Lepidoptera
virus_host_insects2_lepido=virus_host_insects2[grep("Lepidoptera", virus_host_insects2$`host lineage`),]
# subset to Orthoptera
virus_host_insects2_ortho=virus_host_insects2[grep("Orthoptera", virus_host_insects2$`host lineage`),]
# subset to Odonata
virus_host_insects2_odonat=virus_host_insects2[grep("Odonata", virus_host_insects2$`host lineage`),]
# subset to Hemiptera
virus_host_insects2_hemi=virus_host_insects2[grep("Hemiptera", virus_host_insects2$`host lineage`),]
```


```{r}
table(virus_host_insects2$genomic_struture)[c(2,4, 1, 3)]
table(virus_host_insects2_hymeno$genomic_struture)[c(2,4, 1, 3)]
table(virus_host_insects2_coleo$genomic_struture)[c(2,4, 1, 3)]
table(virus_host_insects2_diptera$genomic_struture)[c(2,4, 1, 3)]
table(virus_host_insects2_hemi$genomic_struture)[c(2,4, 1, 3)]
table(virus_host_insects2_hymeno$genomic_struture)[c(2,4, 1, 3)]
table(virus_host_insects2_lepido$genomic_struture)[c(2,4, 1, 3)]
table(virus_host_insects2_odonat$genomic_struture)[c(2,4, 1, 3)]
```



Plot the distribution for major insect orders

```{r, echo=FALSE}
barplot(table(virus_host_insects2$genomic_struture)[c(2,4, 1, 3)], 
        main=paste0("Insects n=", dim(virus_host_insects2)[1]),
        ylab="number of virus species")

barplot(table(virus_host_insects2_hymeno$genomic_struture)[c(2,4, 1, 3)], 
        main=paste0("Hymenoptera n=", dim(virus_host_insects2_hymeno)[1]),
        ylab="number of virus species")

barplot(table(virus_host_insects2_diptera$genomic_struture)[c(2,4, 1, 3)], 
        main=paste0("Diptera n=", dim(virus_host_insects2_diptera)[1]),
        ylab="number of virus species")

barplot(table(virus_host_insects2_coleo$genomic_struture)[c(2,4, 1, 3)], 
        main=paste0("Coleoptera n=", dim(virus_host_insects2_coleo)[1]),
        ylab="number of virus species")

barplot(table(virus_host_insects2_lepido$genomic_struture)[c(2,4, 1, 3)], 
        main=paste0("Lepidoptera n=", dim(virus_host_insects2_lepido)[1]),
        ylab="number of virus species")

barplot(table(virus_host_insects2_ortho$genomic_struture)[c(2,4, 1, 3)], 
        main=paste0("Orthoptera n=", dim(virus_host_insects2_ortho)[1]),
        ylab="number of virus species")

barplot(table(virus_host_insects2_odonat$genomic_struture)[c(2,4, 1, 3)], 
        main=paste0("Odonata n=", dim(virus_host_insects2_odonat)[1]),
        ylab="number of virus species")

barplot(table(virus_host_insects2_hemi$genomic_struture)[c(2,4, 1, 3)], 
        main=paste0("Hemiptera n=", dim(virus_host_insects2_hemi)[1]),
        ylab="number of virus species")

```


Test the difference between all insects and hymeno :
```{r}
chisq.test(rbind(table(virus_host_insects2$genomic_struture)[c(2,4, 1, 3)],
                 table(virus_host_insects2_hymeno$genomic_struture)[c(2,4, 1, 3)]))
```



